---
title: "R Markdown Assignment"
author: "Dwayne"
date: "20/09/2017"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Part 1: Reading in the Data and loading Libraries
``` {r echo=TRUE}
library(dplyr)
library(ggplot2)
df<-tbl_df(read.csv("activity.csv"))
df$date<-as.Date(df$date, format="%Y-%m-%d")
df_filtered<-df[complete.cases(df$steps), ]

```

##Part 2: Histogram of the total number of steps taken each day - 
```{r}
knitr::opts_chunk$set(echo = TRUE)
df_grouped<-group_by(df, by=date)
df_summary <- summarize(df_grouped, Steps=sum(steps))
print(ggplot(df_summary, aes(x=by, y=Steps))+geom_bar(stat="identity")+xlab("Date")+ggtitle("Plot of Number of Steps per Date"))
```

##Part 3: Mean and median number of steps taken each day
```{r}
df_mean_median <- summarize(df_grouped,Mean=mean(steps, na.rm = TRUE), Mdn=median(steps,na.rm = TRUE))
print(ggplot(df_mean_median, aes(x=by, y=Mean))+geom_line(aes(y=Mean,color="Mean"))+geom_line(aes(y=df_mean_median$Mdn,color="Median"))+xlab("Date")+ylab("Mean & Median"))
```

##Part 4: Time series plot of the average number of steps taken
```{r}
df_time <- df_filtered %>% group_by(by=interval) %>% summarize(Steps=mean(steps))
print(ggplot(df_time, aes(x=by, y=Steps))+geom_line()+xlab("Interval")+ylab("Mean Number of Steps"))
```

##Part 5: The 5-minute interval that, on average, contains the maximum number of steps
```{r}
Max_steps<- df_time[which.max(df_time$Steps),1]
```
The 5-minute interval that, on average, contains the maximum number of steps `r Max_steps `.

##Part 6: Code to describe and show a strategy for imputing missing data 
Computing % of observations with missing steps
```{r, echo=FALSE}
missingobservations <- sum(is.na(df$steps))/nrow(df)*100
```
The percentage of observations with missing number of steps is `r missingobservations`%.
We shall use the MICE package to impute the missing data
```{r, echo=TRUE}
library(mice)
imputed_data <- mice(df[,c(1,3)], m=5)
```

Using the 5th set of data of imputed data to complete data:
```{r}
completeData<-complete(imputed_data,5) #using the 5th set of imputed data
completeData<-cbind(completeData,df$date)
```

##Part 7: Histogram of the total number of steps taken each day after missing values are imputed 
```{r}
df_grouped <- group_by(completeData, by=df$date)
df_summary <- summarize(df_grouped, Steps=sum(steps))
print(ggplot(df_summary, aes(x=by, y=Steps))+geom_bar(stat="identity")+xlab("Date")+ggtitle("Plot of Number of Steps per Date"))
```

##Part 8: Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
df_filtered$wDay <- factor((weekdays(df_filtered$date) %in% weekdays1), levels=c(FALSE, TRUE), labels=c('weekend','weekday'))

df_wday_grouped<-df_filtered %>% group_by(interval, wDay) %>% summarise(Mn=mean(steps))
yweekend<-df_wday_grouped$Mn[df_wday_grouped$wDay=="weekend"]
yweekday<-df_wday_grouped$Mn[df_wday_grouped$wDay=="weekday"]

intervals<-unique(df_wday_grouped$interval)
p1<- ggplot(df_wday_grouped, aes(x=intervals, y=yweekend))+geom_line()

yweekend<-df_wday_grouped$Mn[df_wday_grouped$wDay=="weekend"]
yweekday<-df_wday_grouped$Mn[df_wday_grouped$wDay=="weekday"]

par(mfrow=c(2,1))
plot(intervals, yweekend, xlab="Intervals", type = "l",ps=2, ylab="Average Steps", main="Average Steps per Interval during Weekend")
plot(intervals, yweekday, xlab="Intervals",type = "l", ps=2, ylab="Average Steps", main="Average Steps per Interval during Weekday")
```


